Understanding the nanoscale structure of hexagonal phase lyotropic liquid crystal membranes

### 2018-02-23 Ben_Manuscripts/structure_paper/Draft.tex
New figures, updated figures, text edits on draft and supplemental

  A highly selective membrane can be useful for the recovery of valuable
  products from complex aqueous and organic solutions.  For example, flowback
  water (FW) produced during hydraulic fracturing is a complex wastewater full of
  potentially valuable dissolved organic compounds such as acetate
  \cite{dischinger_application_2017}. There is increasing pressure to reuse
  hydraulic fracturing water rather than dispose of it in order to reduce social
  and environmental impacts as well as cost \cite{theodori_hydraulic_2014}.
  Rather fully than dispose of the wastestream generated in the recycling
  process, we can instead use highly selective membranes in order to successfully
  recover useful compounds.

@@

%MRS5: under solution-diffusion, then the free energy to partition into the membrane is important as well, correct?
%BJC: but wouldn't the free energy just be a function of all the factors I list? 
%Maybe I can rephrase it that way: "A solutes ability to enter a given membrane barrier is governed by the free energy to partition into the membrane which is influenced by ..."
  Selective separation by a semipermeable membrane barrier can be broken into two
  steps: barrier entry and transport through the barrier. A solute's ability to
  % BJC: can also come up with a third step: exit. Which I think is mostly chemical potential gradient
  enter a given membrane barrier is decided by its size, shape, charge, polarity
  and how those factors combine to interact with the membrane material. The same
  properties affect the rate at which a solute travels across the barrier. Since
  undesired solutes may still enter the membrane, their diffusion rates relative
  to desired solutes are important
  \cite{gin_polymerized_2008,wijmans_solution-diffusion_1995}. To achieve optimum
  selectivity, it is necessary to design membranes in a way that one can tune the
  relative diffusivities of desired and undesired solutes.

@@

  Crosslinked lyotropic liquid crystal (LLC) membranes may be capable of
  performing highly selective separations. LLCs are amphiphilic molecules that
  have the ability to self-assemble into porous nanostructures
  \cite{smith_ordered_1997} which can be crosslinked to create mechanically
  strong membrane films with pores on the order of 1 nm in diameter
  \cite{zhou_supported_2005}. Unlike commercial NF membranes, LLC membrane pores
  are uniform in size. Since LLC membranes lack a pore size distribution, they
  inherently exhibit high selectivity due to their strict molecular weight
  cut-off (MWCO) ~\cite{zhou_supported_2005}. Additionally, the LLCs chosen to
  make membranes for this application are salts, and therefore lead to Donnan
  exclusion of ions in solution. The membrane gains a net surface charge when
  counterions from head groups that line the pore walls escape into the feed
  solution in order to balance the gradients of concentration and electric
  potential \cite{donnan_theory_1995}.

@@

  \subsubsection{\textit{Radial distribution functions}}
  Closely related to the spatial correlation functions, we explored the pores'
  compositions by measuring the average number densities of various monomer
  components as a function of distance from the pore centers. We looked at the
  average number density of sodium ions, aromatic rings and carbon atoms making
  up the monomer tails. We binned the radial distance of all atoms in each group
  from the pore centers, then normalized by the volume of the annulus defined by
  the bin edges and the z box vector.

@@

  We calculated ionic conductivity using two different methods for robustness.
  The Nernst-Einstein relation, relates the DC ionic conductivity, $\sigma$, to ion
  diffusivity, $D$, concentration, $C$, ion charge, $q$, the Boltzmann constant,
  $k_b$, and temperature, $T$: 
  \begin{equation}
	\sigma = \dfrac{q^2CD}{k_b T} 
	\label{eqn:nernst_einstein}
  \end{equation}
  We measured sodium ion diffusion coefficients by calculating the slope
  of the linear region of the z-direction mean square displacement curve as
  indicated by the Einstein relation \cite{einstein_investigations_1956}. We
  visualized the MSD plot to determine where to begin and end a linear fit. We
  measured ion concentration with respect to the volume of the entire unit cell.

  The second method, termed the Collective Diffusion model, measures the
  movement of the collective variable, Q, which is defined as the amount of
  charge transfer through the system and can be thought to represent the center
  of charge of the system. The conductance, $\gamma$, of the system can be
  calculated as:
  \begin{equation}
	 \gamma = \dfrac{D_Q}{k_b T} 
	\label{eqn:collective_diffusion}
  \end{equation}
  We convert the resulting value to ionic conductivity by multiplying by
  channel length and dividing by the membrane cross sectional area. $D_Q$ is the
  diffusion coefficient of the collective variable Q. It is calculated using the
  Einstein relation. One can access a detailed derivation of the model elsewhere
  \cite{liu_collective_2013}.

@@

  Our simulations best support a model built with 5 monomers per layer based on
  the measured equilibrated pore-to-pore distances. To discern the composition of
  the monomer layers, addressing Question (\ref{point:monomernum}), we ran
  simulations of systems created with 4 - 8 monomers per layer. We built systems
  in both the parallel displaced and sandwiched configurations with layers
  initially spaced 3.7 \AA~apart. We prepared equilibrated configurations
  according to the dry equilibration procedure. All systems are stable after 400
  ns of simulation. In a sense, all systems are at least metastable, addressing
  Question (\ref{point:metastable}), however not all will make physical sense or fit
  the experimental profile that we are trying to match. Figure~\ref{fig:p2p}
  shows the pore spacing for all systems tested. Systems built with 5 monomers in
  each layer equilibrate to a pore spacing that is most consistent with the
  experimental value of 4.12 nm derived from SAXS measurements
  (Figure~\ref{fig:SAXS}). The remainder of this discussion will focus on the
  analysis of systems built with 5 monomers per layer.

@@

  We answer Question (\ref{point:orientation}) by simulating X-ray diffraction
  patterns produced from equilibrated MD trajectories. We tested systems built
  with 5 monomers per layer in the parallel displaced and sandwiched
  configurations. We generated simulated patterns using portions of simulation
  trajectory after equilibration. The patterns for both structures are shown and
  compared to experiment in Figure \ref{fig:XRDsim}.

@@

  %BJC: adjust figure size and alignment -- probably easiest to set figure size in matplotlib
  \begin{figure}
  \begin{subfigure}{0.3\linewidth}
        \centering
        \vspace{-0.2em}
        \includegraphics[width=1.018\linewidth]{offset_rzplot.png}
        \caption{}~\label{fig:rz_offset}
  \end{subfigure}
  \begin{subfigure}{0.3\linewidth}
        \centering
%        \vspace{0.25em}
        \includegraphics[scale=0.29]{WAXS_raw.png}
        \caption{}~\label{fig:raw_waxs}
  \end{subfigure}
  \begin{subfigure}{0.3\linewidth}
        \centering
        \includegraphics[width=\linewidth]{layered_rzplot.png}
        \caption{}~\label{fig:rz_layered}
  \end{subfigure}
  \begin{subfigure}{0.0544\linewidth}
        \centering
        \vspace{-4.00em}
        \includegraphics[width=\linewidth]{colorbar_jet.png}
  \end{subfigure}
  \caption{The simulated XRD pattern generated from the equilibrated trajectory
	  of the parallel displaced configuration (a) exhibits all major reflections
	  present in the experimental WAXS pattern (b). The simulated XRD pattern
	  generated from the equilibrated trajectory in the sandwiched configuration (c) shows
	  all major reflections except R-helix.}
  \label{fig:XRDsim}
  \end{figure}

@@

@@

  The evidence supports the hypothesis that alkane chain tails pack in a
  hexagonal array. Groups of three tails associated with each monomer prefer to
  sit between tails of vertically stacked monomers rather than stacking directly
  on top of them (Figure blah blah blah). This type of packing has been seen
  before in XRD studies of hydrocarbon chains~\cite{small_lateral_1984}.

@@

  \begin{figure}[!htb]
  \centering
  	\begin{subfigure}{\linewidth}
	\centering
		\begin{subfigure}{0.45\textwidth}
        		\centering
        		\includegraphics[width=\linewidth]{offset_tail_packing.png}
        		\caption{}~\label{fig:offset_tails}
		\end{subfigure}
		\begin{subfigure}{0.45\textwidth}
		\centering
	        	\includegraphics[width=\linewidth]{offset_angle_v_I.png}
		        \caption{}~\label{fig:offset_integration}
		\end{subfigure}
	\end{subfigure}
	\begin{subfigure}{\linewidth}
	\centering
		\begin{subfigure}{0.45\textwidth}
	        \centering
		        \includegraphics[width=\linewidth]{angles_traj_layered.png}
		        \caption{}~\label{fig:layered_tails}
		\end{subfigure}
		\begin{subfigure}{0.45\textwidth}
        	\centering
		        \includegraphics[width=\linewidth]{layered_angle_v_I.png}
		        \caption{}~\label{fig:layered_integration}
		\end{subfigure}
	\end{subfigure} 
  \caption{We hypothesize that R-spots is the result of ordered tail packing.
	  Defining the membrane plane to be 0\degree, we
	  measured the angles between each alkane chain tail centroid and its nearest
	  neighbor centroids for the equilibrated parallel displaced (a) and sandwiched
	  (c) configurations. Peaks that appear in each distribution are centered near
	  $\pm~33\degree$. We radially integrated the simulated XRD patterns of the
	  parallel displaced and sandwiched configuration within the region bounding
	  R-alkanes ((b) and (d) respectively).  Peaks appear in the same location as the
	  angle distributions which corroborates our hypothesis. Further, we note that
	  the peaks are strongest in both plots associated with the sandwiched
	  configuration. As shown in Figure \ref{fig:rz_layered}, systems simulated in
	  the sandwiched configuration exhibit more intense R-spots
	  reflections.}~\label{fig:tail_packing}
  \end{figure}

@@

  We learned that layers are well-defined and persistent, answering Question
  (\ref{point:layers}). We established our conclusion by plotting the two
  dimensional pair distribution function, $g(y,z)$ of head group phenyl rings
  using Equation~\ref{eqn:correlation} (Figure \ref{fig:yz_correlation}). In both
  configurations, the phenyl rings stack in the z-direction, along $y=0$, and
  clear layers are formed. The layers in the sandwiched configuration are more
  defined than those in the parallel displaced configuration and are strongly
  correlated with adjacent layers. We also observe correlation of layers between
  pores indicated by the non uniformity in $g(y,z)$ in adjacent columns (along $y
  = \pm 3.65$). Parallel displaced layers display weaker correlations but do not
  decay in their intensity as z increases. We do not see correlation between
  pores, however correlations within pores are strong throughout the length of
  the unit cell since there is no noticeable decay in layer intensity.

@@

  All systems up to this point were built with layers stacked 3.7 \AA~apart.
  When we build systems with layers stacked 5.0 \AA~apart and then let them
  equilibrate, we observe long-term stability of a qualitatively different
  configuration, suggesting that we have found another metastable free energy
  basin, further corroborating Question (\ref{point:metastable}). 
  %The basins are
  %differentiated by their relative degrees of ordering resulting from different
  %spatial arrangements. 
  For reasons we will develop in the next segment of our
  discussion, configurations intially built with layers stacked 5.0 \AA~apart 
  will be grouped into the disordered pore basin, while those built with layers
  initially stacked 3.7 \AA~apart will be grouped into the ordered pore basin.

@@

  The simulated X-ray diffraction patterns indicate further structural
  differences. In the parallel displaced configuration, almost all contrast
  between R-spots and R-alkanes fades (Figure~\ref{fig:offset_disordered_xrd}).
  In the sandwiched configuration, R-spots is weakly present, but in different
  locations, showing higher intensity at the top and bottom of the pattern as
  well as at the intersection of R-alkanes with q$_z$ = 0
  (Figure~\ref{fig:layered_disordered_xrd}).

@@

  \begin{figure}[!hbt]
        \centering
        \begin{subfigure}{0.45\linewidth}
                \centering
                \includegraphics[width=\linewidth]{dbwl.png}
                \caption{}~\label{fig:dbwl_disordered}
        \end{subfigure}%
        \begin{subfigure}{0.45\linewidth}
                \centering
                \includegraphics[width=\linewidth]{p2p2.png}
                \caption{}~\label{fig:p2p_disordered}
        \end{subfigure}
        \begin{subfigure}{0.45\linewidth}
                \centering
                \includegraphics[width=\linewidth]{offset_disordered_rzplot.png}
                \caption{}~\label{fig:offset_disordered_xrd}
        \end{subfigure}%
        \begin{subfigure}{0.45\linewidth}
                \centering
                \includegraphics[width=\linewidth]{layered_disordered_rzplot.png}
                \caption{}~\label{fig:layered_disordered_xrd}
        \end{subfigure}
	\caption{There are distinct differences between the disordered and ordered pore systems. 
		When layers are initially stacked 5\AA~ apart, the distance
		between layers increases (a) and the pore spacing decreases (b) in comparison to
		the ordered pore systems. The simulated
		XRD patterns of disordered pore systems are also appreciably different, most 
		notably in the region bounded by R-alkanes. The disordered pore
		systems started in the parallel displaced configuariton (c) does not show R-spots
		or R-$\pi$ which suggests that head groups are not associating with each other
		and tails are distributed nearly isotropically. The disordered pore system
		started in the sandwiched configuration (d) does not show R-$\pi$ but R-spots 
		is present in different locations because the tails are packed in a different
		motif.}
  \end{figure}

@@

  \begin{figure}
  \begin{subfigure}{1\linewidth}
        \centering
        \begin{subfigure}{0.47\linewidth}
                \centering
                \includegraphics[width=\linewidth]{layered_xy_correlation.png}
                \caption{Sandwiched, Ordered Pore Basin}~\label{fig:sandwich_xy}
        \end{subfigure}%
        \begin{subfigure}{0.47\linewidth}
                \centering
                \includegraphics[width=\linewidth]{disorder_xy_correlation.png}
                \caption{Sandwiched, Disordered Pore Basin}~\label{fig:disorder_sandwich_xy_correlation}
       \end{subfigure}%
       \begin{subfigure}{0.47\linewidth}
                \centering
                \includegraphics[width=\linewidth]{offset_xy_correlation.png}
                \caption{Parallel Displaced, Ordered Pore Basin}~\label{fig:offset_xy}
        \end{subfigure}
        \begin{subfigure}{0.47\linewidth}
                \centering
                \includegraphics[width=\linewidth]{disorder_offset_xy_correlation.png}
                \caption{Parallel Displaced, Disordered Pore Basin}~\label{fig:disorder_offset_xy_correlation}
        \end{subfigure}
  \end{subfigure}
  \caption{$g(x,y)$ for sandwiched and parallel displaced configurations in the
	  ordered and disordered pore basins illustrates the increase in xy plane
	  disorder when initial configuration layer stacking distance is increase from
	  3.7\AA~ to 5\AA~apart.}\label{fig:xy_correlation}
  \end{figure}

@@

  %BJC: Matt thinks this section is necessary

@@

  In order to address Question (\ref{point:poredefinition}), we plotted the
  number densities of heavy atoms in the head group, carbon atoms in the tail
  region and all sodium ions (Figure~\ref{fig:densities}). For the head group
  region, we used the carbon atoms making up the aromatic ring. For the tail
  region we used only carbon atoms of the monomer tails (See Supplemental
  Information for diagram). We average the histograms over at least 50 ns of
  equilibrated trajectory.

@@

  We explored the affect of water on pore structure, addressing Question
  (\ref{point:water}), by preparing parallel displaced and sandwiched
  configurations according to the wet equilibration procedure. There is no
  experimental measurement of trace water concentration in the pores so we tested
  a range of water concentrations from 1 to 5 percent. Our lower bound models a
  system with on average 2 water molecules for each monomer layer.
  % BJC: double check this^ calculation or omit
  Figure~\ref{fig:solvation} shows the simulated diffraction patterns resulting
  from each configuration.

@@

  The system's structure and physical characteristics did not change
  significantly when we applied the cross-linking algorithm to the equilibrated
  parallel-displaced configuration in the ordered pore basin. We simulated
  the cross-linked system in the NPT ensemble for 100 ns. After the system is
  cross-linked, the distance between pores shrinks by 0.4 \AA~and the distance
  between layers increases by 0.04 \AA. All major features are still present in
  the simulated XRD patterns, however at lower intensities
  (Fig.~\ref{fig:rzplot_xlink}). We calculated the ionic conductivity using the
  Nernst-Einstein relation and found that it is lower in the cross-linked system
  (Fig.~\ref{fig:IC_xlink}).

>>>>

  A highly selective membrane would be useful for the recovery of valuable
  products from complex aqueous and organic solutions.  For example, flowback
  water (FW) produced during hydraulic fracturing is a complex wastewater full of
  potentially valuable dissolved organic compounds such as acetate
  \cite{dischinger_application_2017}. There is increasing pressure to reuse
  hydraulic fracturing water rather than dispose of it in order to reduce social
  and environmental impacts as well as cost \cite{theodori_hydraulic_2014}.
  Rather fully than dispose of the wastestream generated in the recycling
  process, we can instead use highly selective membranes in order to successfully
  recover useful compounds.

@@

%MRS5: under solution-diffusion, then the free energy to partition into the membrane is important as well, correct?
%  Selective separation by a semipermeable membrane barrier can be broken into two
%  steps: barrier entry and transport through the barrier. A solute's ability to
  % BJC: can also come up with a third step: exit. Which I think is mostly chemical potential gradient
%  enter a given membrane barrier is decided by its size, shape, charge, polarity
%  and how those factors combine to interact with the membrane material. The same
%  properties affect the rate at which a solute travels across the barrier. Since
%  undesired solutes may still enter the membrane, their diffusion rates relative
%  to desired solutes are important
%  \cite{gin_polymerized_2008,wijmans_solution-diffusion_1995}. To achieve optimum
%  selectivity, it is necessary to design membranes in a way that one can tune the
%  relative diffusivities of desired and undesired solutes.

  Selective separation by a semipermeable membrane barrier is a function of the
  geometric and chemical interactions of solutes with the membrane material. A
  molecule's size, shape, charge and polarity combine to determine the degree to
  which a solute partitions into a membrane and how fast it travels through the
  membrane. To separate a component from a mixture, one must understand how to
  design membranes in order to tune the relative transport rates of desired and
  undesired solutes\cite{gin_polymerized_2008,wijmans_solution-diffusion_1995}.

@@

  Crosslinked lyotropic liquid crystal (LLC) membranes may be capable of
  performing highly selective separations. LLCs are amphiphilic molecules that
  have the ability to self-assemble into porous nanostructures
  \cite{smith_ordered_1997} which can be crosslinked to create mechanically
  strong membrane films with pores on the order of 1 nm in diameter
  \cite{zhou_supported_2005}. Unlike commercial NF membranes, LLC membrane pores
  are uniform in size. Since LLC membranes lack a pore size distribution, they
  inherently exhibit high selectivity due to their strict molecular weight
  cut-off (MWCO) ~\cite{zhou_supported_2005}. Additionally, the LLC monomers
  chosen to make membranes for this application are salts, and therefore lead to
  Donnan exclusion of ions in solution. The membrane gains a net surface charge
  when counterions from head groups that line the pore walls escape into the feed
  solution in order to balance the gradients of concentration and electric
  potential \cite{donnan_theory_1995}.

@@

  \subsubsection{\textit{Radial distribution functions}}
  Closely related to the spatial correlation functions, we explored the pores'
  compositions by measuring the average number densities of various monomer
  components as a function of distance from the pore centers. We looked at the
  average number density of sodium ions, aromatic rings and carbon atoms making
  up the monomer tails. We binned the radial distance of all atoms in each group
  from the pore centers, then normalized by the volume of the annulus defined by
  the bin edges and the z box vector (See supplemental information for a 
  diagram).

@@

  We calculated ionic conductivity using two different methods for robustness.
  The Nernst-Einstein relation, relates the DC ionic conductivity, $\sigma$, to ion
  diffusivity, $D$, concentration, $C$, ion charge, $q$, the Boltzmann constant,
  $k_b$, and temperature, $T$: 
  \begin{equation}
	\sigma = \dfrac{q^2CD}{k_b T} 
	\label{eqn:nernst_einstein}
  \end{equation}
  We measured sodium ion diffusion coefficients by calculating the slope
  of the linear region of the z-direction mean square displacement curve as
  indicated by the Einstein relation \cite{einstein_investigations_1956}. We
  visualized the MSD plot to determine where to begin and end a linear fit. We
  measured ion concentration with respect to the volume of the entire unit cell. 
  The second method, termed the Collective Diffusion model, measures the
  movement of the collective variable, Q, which is defined as the amount of
  charge transfer through the system and can be thought to represent the center
  of charge of the system. The conductance, $\gamma$, of the system can be
  calculated as:
  \begin{equation}
	 \gamma = \dfrac{D_Q}{k_b T} 
	\label{eqn:collective_diffusion}
  \end{equation}
  We convert the resulting value to ionic conductivity by multiplying by
  channel length and dividing by the membrane cross sectional area. $D_Q$ is the
  diffusion coefficient of the collective variable Q. It is calculated using the
  Einstein relation. One can access a detailed derivation of the model elsewhere
  \cite{liu_collective_2013}.

@@

  Our simulations best support a model built with 5 monomers per layer based on
  the measured equilibrated pore-to-pore distances. To discern the composition of
  the monomer layers, addressing question (\ref{point:monomernum}), we ran
  simulations of systems created with 4 - 8 monomers per layer. We built systems
  in both the parallel displaced and sandwiched configurations with layers
  initially spaced 3.7 \AA~apart. We prepared equilibrated configurations
  according to the dry equilibration procedure. All systems are stable after 400
  ns of simulation. In a sense, all systems are at least metastable, addressing
  question (\ref{point:metastable}), however not all will make physical sense or fit
  the experimental profile that we are trying to match. Figure~\ref{fig:p2p}
  shows the pore spacing for all systems tested. Systems built with 5 monomers in
  each layer equilibrate to a pore spacing that is most consistent with the
  experimental value of 4.12 nm derived from SAXS measurements
  (Figure~\ref{fig:SAXS}). The remainder of this discussion will focus on the
  analysis of systems built with 5 monomers per layer.

@@

  We answer question (\ref{point:orientation}) by simulating X-ray diffraction
  patterns produced from equilibrated MD trajectories. We tested systems built
  with 5 monomers per layer in the parallel displaced and sandwiched
  configurations. We generated simulated patterns using portions of simulation
  trajectory after equilibration. The patterns for both structures are shown and
  compared to experiment in Figure \ref{fig:XRDsim}.

@@

  %BJC: adjust figure size and alignment -- probably easiest to set figure size in matplotlib
  \begin{figure}
  \begin{subfigure}{0.3\linewidth}
        \centering
        \vspace{-0.2em}
        \includegraphics[width=1.1\linewidth,trim={1cm 0 1.3cm 0},clip]{offset_rzplot.png}
        \caption{}~\label{fig:rz_offset}
  \end{subfigure}
  \begin{subfigure}{0.3\linewidth}
        \centering
%        \vspace{0.25em}
        \includegraphics[scale=0.285]{WAXS_raw.png}
        \caption{}~\label{fig:raw_waxs}
  \end{subfigure}
  \begin{subfigure}{0.3\linewidth}
        \centering
        \includegraphics[width=1.1\linewidth,trim={1cm 0 1.3cm 0},clip]{layered_rzplot.png}
        \caption{}~\label{fig:rz_layered}
  \end{subfigure}
  \begin{subfigure}{0.0544\linewidth}
        \centering
        \vspace{-4.00em}
        \includegraphics[width=\linewidth]{colorbar_seismic.png}
  \end{subfigure}
  \caption{The simulated XRD pattern generated from the equilibrated trajectory
	  of the parallel displaced configuration (a) exhibits all major reflections
	  present in the experimental WAXS pattern (b). The simulated XRD pattern
	  generated from the equilibrated trajectory in the sandwiched configuration (c) shows
	  all major reflections except R-helix.}
  \label{fig:XRDsim}
  \end{figure}

@@

  In addition to R-$\pi$ being shifted to lower $\mathbf{q}$ values, a few
  factors combine to make the simulated R-$\pi$ reflection visibly more intense
  than in the experimental pattern. The simplest reason is that R-$\pi$ and
  R-alkanes add together to boost the overall signal at interesecting values of
  $\mathbf{q}$. In addition, we observe a much higher maximum intensity within
  R-$\pi$. The maximum intensity within R-$\pi$ of the normalized experimental
  pattern is 3.1, while the maximum intensity within R-$\pi$ of the normalized
  parallel displaced simulated pattern is 32.9. While direct numerical
  comparisons between simulation and experiment aren't necessarily precise due to
  underlying assumptions, an order of magnitude difference suggests some
  significance. We believe that we see such a high maximum for two reasons: (1)
  We are simulating a perfectly aligned system so all X-rays reflected along the
  pore axis would add constructively. The crystalline domains in the real system
  are not perfectly aligned (See azimuthal distribution in supplemental
  information) which would lead to an overall decrease in the intensity of
  R-$\pi$. (2) Our system is more ordered with high correlation between layers.
  We will come back to this point further in ensuing discussions.

  There is a grid-like pattern near $\mathbf{q}=0$ which appears because the
  simulated system shows crystalline character. The grid features along nonzero
  values of $q_z$ line up with reflections at $q_z=0$ which indicates positional
  correlations between columns in the z-direction. This type of correlation is
  expected for a crystalline material. If our system met the definition of a
  columnar liquid crystalline system, we should observe only weak short-range
  z-directional correlation within columns, with no correlation between
  columns\cite{chaikin_principles_1995}.

@@

  The evidence supports the hypothesis that alkane chain tails pack in a
  hexagonal array. Groups of three tails associated with each monomer prefer to
  sit between tails of vertically stacked monomers rather than stacking directly
  on top of them (Figure~\ref{fig:hex_tails}). This type of packing has been seen
  before in XRD studies of hydrocarbon chains~\cite{small_lateral_1984}.

@@

  \begin{figure}[!htb]
  \centering
  	\begin{subfigure}{\linewidth}
	\centering
		\begin{subfigure}{0.45\textwidth}
        		\centering
        		\includegraphics[width=\linewidth]{offset_tail_packing.png}
        		\caption{}~\label{fig:offset_tails}
		\end{subfigure}
		\begin{subfigure}{0.45\textwidth}
		\centering
	        	\includegraphics[width=\linewidth]{offset_angle_v_I.png}
		        \caption{}~\label{fig:offset_integration}
		\end{subfigure}
	\end{subfigure}
	\begin{subfigure}{\linewidth}
	\centering
		\begin{subfigure}{0.45\textwidth}
	        \centering
		        \includegraphics[width=\linewidth]{angles_traj_layered.png}
		        \caption{}~\label{fig:layered_tails}
		\end{subfigure}
		\begin{subfigure}{0.45\textwidth}
        	\centering
		        \includegraphics[width=\linewidth]{layered_angle_v_I.png}
		        \caption{}~\label{fig:layered_integration}
		\end{subfigure}
	\end{subfigure} 
  \caption{We hypothesize that R-spots is the result of ordered tail packing.
	  Defining the membrane plane to be 0\degree, we measured the angles between each
	  alkane chain tail centroid and its nearest neighbor centroids for the
	  equilibrated parallel displaced (a) and sandwiched (c) configurations. Peaks
	  that appear in each distribution are centered near $\pm~33\degree$. We radially
	  integrated the simulated XRD patterns of the parallel displaced and sandwiched
	  configuration within the region bounding R-alkanes ((b) and (d) respectively).
	  Peaks appear in the same location as the angle distributions which corroborates
	  our hypothesis. Further, we note that the peaks are strongest in both plots
	  associated with the sandwiched configuration. As shown in Figure
	  \ref{fig:rz_layered}, systems simulated in the sandwiched configuration exhibit
	  more intense R-spots reflections. Note that plots of integrated intensity are
	  normalized so the average alkane intensity (excluding reflections $\pm
	  30\degree$ from the $q_z$ axis) is set to 1.  Plots are cut off for intensities
	  greater than 2 so peaks are more visible.}~\label{fig:tail_packing}
  \end{figure}

%BJC: not sure if this figure is necessary
  \begin{figure}
	\centering
	\includegraphics[width=0.5\linewidth]{hexagonal_tails_alt2.png} 
	%BJC: I can give the circle some depth if that makes it clearer
	\caption{Monomer tails prefer to pack in a hexagonal array. Each circle
		represents a monomer tail as seen if viewed by looking from the end of the tail
		to the pore center. Each color represents distinct monomers from adjacent
		layers.}~\label{fig:hex_tails}
  \end{figure}

@@

  We learned that layers are well-defined and persistent, answering question
  (\ref{point:layers}). We established our conclusion by plotting the two
  dimensional pair distribution function, $g(y,z)$ of head group phenyl rings
  using Equation~\ref{eqn:correlation} (Figure \ref{fig:yz_correlation}). In both
  configurations, the phenyl rings stack in the z-direction, along $y=0$, and
  clear layers are formed. The layers in the sandwiched configuration are more
  defined than those in the parallel displaced configuration and are strongly
  correlated with adjacent layers. We also observe correlation of layers between
  pores indicated by the non uniformity in $g(y,z)$ in adjacent columns (along $y
  = \pm 3.65$). Parallel displaced layers display weaker correlations but do not
  decay in their intensity as z increases. We do not see correlation between
  pores, however correlations within pores are strong throughout the length of
  the unit cell since there is no noticeable decay in layer intensity.

@@

  All systems up to this point were built with layers stacked 3.7 \AA~apart.
  When we build systems with layers stacked 5.0 \AA~apart and then let them
  equilibrate, we observe long-term stability of a qualitatively different
  configuration, suggesting that we have found another metastable free energy
  basin, further corroborating question (\ref{point:metastable}). 
  %The basins are
  %differentiated by their relative degrees of ordering resulting from different
  %spatial arrangements. 
  For reasons we will develop in the next segment of our
  discussion, configurations intially built with layers stacked 5.0 \AA~apart 
  will be grouped into the disordered pore basin, while those built with layers
  initially stacked 3.7 \AA~apart will be grouped into the ordered pore basin.

@@

  The simulated X-ray diffraction patterns indicate further structural
  differences. In the parallel displaced configuration, almost all contrast
  between R-spots and R-alkanes fades (Figure~\ref{fig:offset_disordered_xrd}).
  In the sandwiched configuration, R-spots is present, but there are only two
  spots located at the intersection of $q_z=0$ with R-alkanes.
  (Figure~\ref{fig:layered_disordered_xrd}).

@@

  \begin{figure}[!hbt]
        \centering
        \begin{subfigure}{0.45\linewidth}
                \centering
                \includegraphics[width=\linewidth]{dbwl.png}
                \caption{}~\label{fig:dbwl_disordered}
        \end{subfigure}%
        \begin{subfigure}{0.45\linewidth}
                \centering
                \includegraphics[width=\linewidth]{p2p2.png}
                \caption{}~\label{fig:p2p_disordered}
        \end{subfigure}
        \begin{subfigure}{0.45\linewidth}
                \centering
                \includegraphics[width=\linewidth]{disorder_offset_rzplot.png}
                \caption{}~\label{fig:offset_disordered_xrd}
        \end{subfigure}%
        \begin{subfigure}{0.45\linewidth}
                \centering
                \includegraphics[width=\linewidth]{disorder_rzplot.png}
                \caption{}~\label{fig:layered_disordered_xrd}
        \end{subfigure}
	\caption{There are distinct differences between the disordered and ordered pore systems. 
		When layers are initially stacked 5\AA~ apart, the distance
		between layers increases (a) and the pore spacing decreases (b) in comparison to
		the ordered pore systems. The simulated
		XRD patterns of disordered pore systems are also appreciably different, most 
		notably in the region bounded by R-alkanes. The disordered pore
		systems started in the parallel displaced configuariton (c) does not show R-spots
		or R-$\pi$ which suggests that head groups are not associating with each other
		and tails are distributed nearly isotropically. The disordered pore system
		started in the sandwiched configuration (d) does not show R-$\pi$ but R-spots 
		is present in different locations because the tails are packed in a different
		motif.}
  \end{figure}

@@

  \begin{figure}
  \begin{subfigure}{1\linewidth}
        \centering
        \begin{subfigure}{0.47\linewidth}
                \centering
                \includegraphics[width=\linewidth]{layered_xy_correlation.png}
                \caption{Sandwiched, Ordered Pore Basin}~\label{fig:sandwich_xy}
        \end{subfigure}
        \begin{subfigure}{0.47\linewidth}
                \centering
                \includegraphics[width=\linewidth]{disorder_xy_correlation.png}
                \caption{Sandwiched, Disordered Pore Basin}~\label{fig:disorder_sandwich_xy_correlation}
       \end{subfigure}
       \begin{subfigure}{0.47\linewidth}
                \centering
                \includegraphics[width=\linewidth]{offset_xy_correlation.png}
                \caption{Parallel Displaced, Ordered Pore Basin}~\label{fig:offset_xy}
        \end{subfigure}
        \begin{subfigure}{0.47\linewidth}
                \centering
                \includegraphics[width=\linewidth]{disorder_offset_xy_correlation.png}
                \caption{Parallel Displaced, Disordered Pore Basin}~\label{fig:disorder_offset_xy_correlation}
        \end{subfigure}
  \end{subfigure}
  \caption{$g(x,y)$ for sandwiched and parallel displaced configurations in the
	  ordered and disordered pore basins illustrates the increase in xy plane
	  disorder when initial configuration layer stacking distance is increase from
	  3.7 \AA~to 5 \AA~apart.}\label{fig:xy_correlation}
  \end{figure}

@@

@@

  In order to address question (\ref{point:poredefinition}), we plotted the
  number densities of heavy atoms in the head group, carbon atoms in the tail
  region and all sodium ions (Figure~\ref{fig:densities}). For the head group
  region, we used the carbon atoms making up the aromatic ring. For the tail
  region we used only carbon atoms of the monomer tails (See Supplemental
  Information for diagram). We average the histograms over at least 50 ns of
  equilibrated trajectory.

@@

  % BJC: I'm not sure this section is necessary anymore. The original point was to 
  % see if we get more structure. It's not clear this is the right way to introduce
  % water into the system. I think it confuses things.

  We explored the affect of water on pore structure, addressing question
  (\ref{point:water}), by preparing parallel displaced and sandwiched
  configurations according to the wet equilibration procedure. There is no
  experimental measurement of trace water concentration in the pores so we tested
  a range of water concentrations from 1 to 5 percent. Our lower bound models a
  system with on average 2 water molecules for each monomer layer.
  % BJC: double check this^ calculation or omit
  Figure~\ref{fig:solvation} shows the simulated diffraction patterns resulting
  from each configuration.

@@

  The system's structure and physical characteristics did not change
  significantly when we applied the cross-linking algorithm to the equilibrated
  parallel-displaced configuration in the ordered pore basin. We simulated
  the cross-linked system in the NPT ensemble for 100 ns. After the system is
  cross-linked, the distance between pores shrinks by 0.4 \AA~and the distance
  between layers increases by 0.04 \AA. All major features are still present in
  the simulated XRD patterns, however at lower intensities
  (Fig.~\ref{fig:rzplot_xlink}). We calculated the ionic conductivity using the
  Nernst-Einstein relation and found that it is lower in the cross-linked system
  (Fig.~\ref{fig:IC_xlink}).
  %BJC: The current XRD pattern is from a 100 ns simulation. I am running it out
  % longer to get more statistics

  \begin{figure}
  \centering
  \begin{subfigure}{0.45\textwidth}
	\centering
	\includegraphics[width=\textwidth]{rzplot_xlink.png}
	\caption{}~\label{fig:rzplot_xlink}
  \end{subfigure}
  \begin{subfigure}{0.45\textwidth}
	\centering
	\includegraphics[width=\textwidth]{IC_xlink.png}
	\caption{}~\label{fig:IC_xlink}
  \end{subfigure}
  \begin{subfigure}{0.45\textwidth}
	\centering
	\includegraphics[width=\textwidth]{dbwl_xlink.png}
	\caption{}~\label{fig:dbwl_xlink}
  \end{subfigure}
  \begin{subfigure}{0.45\textwidth}
	\centering
	\includegraphics[width=\textwidth]{p2p_xlink.png}
	\caption{}~\label{fig:p2p_xlink}
  \end{subfigure}
  \caption{Applying our simulated crosslinking mechanism to an equilibrium
	  configuration causes slight changes to the system's physical and structural
	  properties. (a) Reflections produced by the cross-linked configuration are
	  faded relative to the uncross-linked system. (b) The ionic conductivity is
	  smaller relative to the uncross-linked system, but still much larger than the
	  experimental value. When the system is cross-linked, the distance between
	  layers increases (c) and the pore spacing decreases (d)}~\label{fig:xlink}
  \end{figure}
